A Method for the Spectroscopic Inference of Stellar Parameters

### 2014-07-30 tex/paper.tex
Ready for run on Odyssey

Here we opt to take the computationally expedient approach that employs a library of model spectra, 
$\flam(\{\vt_{\ast}\}^{\rm grid})$, where $\vt_{\ast} = [T_{\rm eff}, \,\, \log{g}, \,\, Z]$ 
(in practice, the metallicity $Z$ is often parameterized by [Fe/H]).  However, it is worth noting 
that the techniques we will develop are applicable to {\it any} reasonable ``front-end" that 
generates a model spectrum.  In our adopted approach, the model spectrum for an arbitrary 
$\vt_{\ast}$ must be interpolated from among the spectral library, 
\begin{equation}
\flam(\{\vt_{\ast}\}^{\rm grid}) \leadsto \flam(\vt_{\ast}),
\label{eqn:interp}
\end{equation}
where we assign the symbol $\leadsto$ as an interpolation operator.  The multi-dimensional 
interpolation in Eq.~\ref{eqn:interp} needs to be performed many times, so computational efficiency 
is critical.  In practice, a simple tri-linear interpolation is suitably fast, but introduces an 
undesireable level of inaccuracy (particularly in the $Z$ dimension).  The interpolation quality 
can be empirically estimated by performing the operation in Eq.~\ref{eqn:interp} across a 
calculated location in $\{\vt_{\ast}\}^{\rm grid}$, and then comparing the interpolated spectrum 
with the corresponding library spectrum.  After an extensive exploration of such calculations 
\citep[see also][]{husser12}, we concluded that the best combination of speed and accuracy can be 
achieved by pre-computing a {\it refined} spectral library using a cubic spline interpolation with 
a $\{\vt_{\ast}\}^{\rm grid}$ spacing of [20\,K, 0.1\,dex, 0.1\,dex], and then performing 
tri-linear interpolation over that refined grid.  Overall, this interpolation technique is found to 
be accurate within a few percent per high resolution model pixel.  Ideally, this pre-interpolation 
could be avoided if the spectral library was computed over a refined grid (with a substantial 
up-front computational investment); but for the time being, we can empirically propagate these 
interpolation uncertainties into the likelihood calculations (as will be described in 
Section~\ref{subsec:likelihood}).

@@

We can further divide $\vt_{\rm obs}$ into those parameters that impact the model primarily in the 
spectral or flux dimensions; $\vt_{\rm obs} = [\vt_{\rm obs, v}, \,\, \vt_{\rm obs, f}]$.  For the 
former, we consider three kernels that contribute to the observed line-of-sight velocity 
distribution function, $\varphi_v$.  The first, $\mathcal{F}_v^{\rm inst}$, treats the instrumental 
spectral broadening.  For illustrative purposes we assume $\mathcal{F}_v^{\rm inst}$ is a Gaussian 
with a mean at $v = 0$ and a constant width $\sigma_v$ at all $\lambda$, although more 
sophisticated forms could be adopted.  The second, $\mathcal{F}_v^{\rm rot}$, characterizes the 
broadening induced by (projected) stellar rotation, parameterized by $v\sin{i}$ as described by 
\citet[][his Eq.~18.14]{gray08}.  And the third, $\mathcal{F}_v^{\rm dop} = \delta(v-v_z)$, 
incorporates the radial velocity through a Doppler shift.  The model spectrum is modified by the 
parameters $\vt_{{\rm obs}, v} = [\sigma_v, \,\, v\sin{i}, \,\, v_z]$ through these kernels, using 
a convolution in velocity-space,\footnote{In practice, these convolutions are performed as 
multiplications in Fourier-space to better preserve spectral information \citep[cf.,][]{tonry79}; 
the mathematical formalism is presented here for clarity.}
\begin{equation}
\flam(\vt_{\ast}, \vt_{{\rm obs}, v}) = \flam(\vt_{\ast}) \otimes \varphi_v = \flam(\vt_{\ast}) \otimes \mathcal{F}_v^{\rm inst} \otimes \mathcal{F}_v^{\rm rot} \otimes \mathcal{F}_v^{\rm dop},
\label{eqn:broadening}
\end{equation}
and then re-sampled onto the discrete wavelengths corresponding to each data pixel, 
\begin{equation}
\flam(\vt_{\ast}, \vt_{{\rm obs}, v}) \mapsto \vM(\vt_{\ast}, \vt_{{\rm obs}, v}),
\label{eqn:resampling}
\end{equation}
where the $\mapsto$ symbol denotes a re-sampling operator that maps the model spectrum onto the 
$N_{\rm pix}$-element model vector $\vM$ (and $N_{\rm pix}$ is the number of pixels in the 
spectrum).  For reference, Figure \ref{fig:broadening} shows a graphical representation of these 
post-processing steps.

@@

The inference problem as framed here has a natural hierarchical structure between the collections 
of ``interesting" parameters, $\vT = [\vt_{\ast}, \vt_{\rm obs}]$, and ``nuisance" hyperparameters, 
$\vP = [\cheb, \cov]$.  To explore the posterior probability distribution of the model conditioned 
on an observed spectrum in this case, it is convenient to use MCMC simulations with a blocked Gibbs 
sampler coupled to the Metropolis-Hastings algorithm.  In essence, this procedure works by sampling 
in a subset of parameters with Metropolis-Hastings conditioned on the current (fixed) values of the 
other parameters; after each iteration, the Gibbs sampler then updates the values of that parameter 
subset and then cycles through all the different parameter subsets in the same way (for a more 
nuanced mathematical description of the process, see Chapter 11 of \citealt{gelman13}).  The 
step-by-step prescription is as follows, where we denote the $i^{\rm th}$ iteration of the Gibbs 
sampler as a superscript index (not to be confused with the $i^{\rm th}$ pixel as in our previous 
notation):
\begin{enumerate}
\item Initialize the parameters.  One might set $\vT^0$ based on estimates in the literature or 
scaling behaviors, and make simple assumptions about $\vP^0$.  Here, we set the Chebyshev 
polynomial coefficients ($\cheb^0$) to be flat ($c_0 = 1$,  $c_{>0} = 0$ in all spectral orders) 
and assume only the trivial noise spectrum contributes to the covariance matrix (all elements of 
$\cov^0$ are zero).

@@

\item Decide whether to instantiate any additional local (non-stationary) covariance kernels.  Here 
we adopt an iterative thresholding criterion: if a local patch of the residual signal is at least 
$3\times$ the amplitude of the global covariance kernel (i.e., where $\vR \ge 3 \, a_g^i$), this 
will trigger the addition of a local kernel.

\item Now sample in the local covariance hyperparameters $\vp_{{\mathsf C},k}$ using one 
iteration of the Metropolis-Hastings algorithm to compute the posterior probabilities
with $\vT^i$, $\cheb^i$, and $\vp_{{\mathsf C}, g}^i$ fixed.  Update the local covariance 
hyperparameters $\vp_{{\mathsf C}, k}^{i-1} \rightarrow \vp_{{\mathsf C}, k}^i$. If the stellar parameters have moved to a region of parameter space where the local kernel is no longer required, then the amplitude of the kernel will be driven towards $0$ and the kernel would have a negligible contribution to the covariance matrix $\vC$. 
\end{enumerate}

@@

\todo{What should we do with the commented-out sections?}
%A major advantage of using the MCMC algorithm to explore the multidimensional probability space is that it provides numerical samples in each dimension. Therefore, marginalizing out a parameter (i.e., numerically integrating over a dimension in probability space) is as simple as combining all of the samples in this dimension. This enables us to present a posterior of the stellar parameters $\vT$ (Figure~\ref{fig:stellar_posterior}) which has been marginalized over all of the nuisance hyperparameters. This posterior is the final estimate of the stellar parameters which \emph{incorporates} any inherent uncertainty due to model mismatch (via the covariance hyperparameters) and flux-calibration (via the Chebyshev polynomials).

%The benefit of including nuisance parameters and then marginalizing over them is that we can self-consistently model the uncertainty inherent to any spectrum while naturally capturing any degeneracy between the model parameters. If stellar parameters are estimated using a method which ignores these nuisance parameters (e.g., by-eye fitting) and the astronomer arbitrarily inflates the parameter uncertainties to reflect intuition, the degeneracy between parameters is artificially destroyed.

%When iteratively fitting a spectrum (see \S\ref{subsec:MCMC}), we continue to add line covariance kernels until we have covered all of the high amplitude residuals. Typically, we will add line kernels until all residuals greater than three times the amplitude of the global covariance kernel are covered.  As demonstrated by the Gibbs sampler routine, we explicitly fit for the
%hyperparameters of the covariance kernels in parallel with the
%stellar parameters. While including these extra parameters does increase the
%dimensionality and complexity of our model, they do provide several advantages.
%\emph{A priori} we do not know which regions of the spectrum are improperly
%modeled. The covariance hyperparameters provide a framework to identify these
%regions iteratively and in a self-consistent manner.

%The traditional manner of dealing with spectral mismatch is to simply mask out the regions of the spectrum which do not agree to within a certain tolerance.  Rather than arbitrarily excluding regions of the spectrum from the fit, these regions should instead be incorporated into in the fit with the appropriate weight. A model that includes covariance is far more likely than forcing the synthetic spectrum to fit perfectly, and far more flexible than arbitrarily masking regions which do not fit. The fitting procedure allows these weights to be determined self-consistently, such that lines which are slightly wrong can still bring information to bear on the stellar parameters. In fact, for some types of stars with difficult-to-model spectra, it may be possible that the lines that are incorrect by a small amount actually provide the \emph{most information} about the stellar parameters, precisely because these lines are the most sensitive to stellar structure and consequentially are the most difficult to model correctly.

%Another powerful benefit of including the covariance hyperparameters is the result of their ability to quantify and account for model-data mismatch. When a data spectrum is fit with a high-quality spectral library, the pixel residuals are less likely to be correlated, and the covariance structure will have a smaller amplitude and correlation length. This means that the inference on the stellar parameters will be more precise, in the same way that high quality data allows a more precise result. However, if the systematic mismatch between the data and model is large, then the hyperparameters will be larger and the precision of the stellar parameters will naturally inflate to respect the quality of model-data fit. The benefits of using covariance hyperparameters also extend beyond the use case of fitting a single stellar spectrum. If we fit many stars with the same set of synthetic models, we can use the structure of the covariance matrix to improve the models themselves (see \S\ref{subsec:learning}).

%\subsection{Applications \comm{untouched}}
%\label{subsec:learning}
%By cataloguing the covariance structure of the residuals, especially those generated from strong spectral line mismatch, we collect valuable information about the quality of the synthetic spectra. After fitting many stars, the accumulated knowledge of the data-model mismatch can be used when fitting a new star. The previous structure of the covariance matrix allows us to set priors on what the covariance in certain regions of the spectrum should be, which will speed convergence for this new star. Additionally, after fitting several stars, the average value of the covariance matrix will inform us about the quality of specific spectral lines in the synthetic models.

%Linking the covariance matrices of stars could be done serially, where the aggregate covariance matrix of all previous fits is used as a prior for the current iteration. Or, the covariance matrices could instead be linked hierarchically, in that the parameters describing the depth and width of a line residual for a specific star $a_1$ and $\sigma_1$, are modeled as coming from of a population of possible depths and widths for a given spectral type. Each stellar spectrum will have a slightly different realization of a spectral line, which will have some scatter about the average residual height. Linking the covariance matrix between spectra of similar stars allows us to grow more confident in our assessment that certain synthetic spectral lines are indeed outliers and should be appropriately down-weighted. In turn, as we become more certain of the weights, the stellar posteriors will become narrower and make our estimates of the stellar parameters more precise. This ability of the model to mutually inform sets of parameters is one of the major advantages of hierarchical Bayesian analysis \citep{kruschke10}.

%Once determined, this average covariance matrix could be delivered to the communities that created the synthetic libraries, which would enable them to rapidly pinpoint and correct non-physical lines in the synthetic models. Alternatively, we could correct the models ourselves by using the chain of logic and mathematical post-processing that we used to created the synthetic model spectrum to reverse-engineer what the behavior of the raw synthetic spectrum \emph{should} be, at the raw $R \gtrsim$~100,000 resolution. This fundamental application of machine learning would enable us to create our own library of data driven, semi-empirical stellar models. Rather than simply assembling an empirical spectral library using only real stellar spectra, this combined approach is more powerful because the synthetic stellar atmospheres provide an actual anchor point of fundamental stellar parameters tied down by the laws of stellar physics.

@@

\subsection{Datasets and historical analysis}
% Paragraphs as follows:
% Introduction to the two types of stars, in order to demonstrate the flexibility.
To demonstrate the flexibility of our probabilistic framework, we model spectra of two different stars acquired with different instruments. The first data set is a high resolution optical spectrum of WASP-14, an F5 star \citep{joshi09}. The second data set is a moderate resolution mid-infrared spectrum of Gl~51, a M5 star \citep{rojas-ayala12}.

% Data description of WASP-14
Our spectrum of WASP-14 was obtained with the Tillinghast Reflector Echelle Spectrograph (\emph{TRES}, \citealt{mink11}) on the Fred Laurence Whipple Observatory 1.5m telescope on June 14th, 2009. \emph{TRES} delivers a 51 echelle spectrum spanning the full optical spectrum (3860--9100\AA) at a resolution of $R \sim$44,000. \citet{torres12} originally acquired the spectrum of WASP-14 to study the stellar properties of transiting exoplanet hosts. The raw spectrum was reduced using the well-tested \emph{TRES} pipeline written by L.~Buchhave. In the \ion{Mg}{0}b echelle order ($\sim$5100\AA), the spectrum has a signal to noise of 150 per resolution element.

% What other people get when they fit this star. How does our inference compare?
\citet{torres12} measured the stellar properties of WASP-14 using several different methods including MOOG, SME, and SPC. The MOOG and SME methods are not directly comparable to our technique because they measure individual line strengths. However, because SPC utilizes a large swath of the spectrum and a model spectrum, we can use the same spectral library and make the same assumptions. The SPC analysis within \citet{torres12} focused on a region of spectrum from 5050--5360\AA\ and used many methods. Additionally, there are systematics introduced by metallicity assumptions and radiative transfer. Using this same spectral library as SPC, we recover the same parameters. When we fit this spectrum using the same assumptions, (amounting to a $\delta$-function prior on $\log g$). This demonstrates sensitivity to a prior as well as the fact that surface gravity estimates from spectroscopy may be biased against fits derived from planetary transits. A Bayesian approach allows you to explore sensitivity to prior assumptions. Letting the covariance float, we achieve comparable estimates. See Table~\ref{table:tests} for our results.

% Data description of Gl 51
The spectrum of Gl~51 was obtained with the \emph{SPEX} instrument \citep{rayner03} on the IRTF. The spectrum is a moderate resolution (R $\sim$ 2,000), mid-infrared spectrum from 0.7--5.5$\mu$m at signal to noise $> 200$. Gl~51 is a M5 star from the IRTF spectral library \citep{cushing05, rayner09}.

% What other people get when they fit this star. How does our inference compare?
To determine the stellar parameters of Gl~51, \citep{rojas-ayala12} use the 2.205$\mu$m \ion{Na}{1} doublet and 2.263$\mu$m \ion{Ca}{1} triplet and their \HK\ index, defined using three passbands: 2.070--2.090$\mu$m, 2.235--2.255$\mu$m, and 2.360--2.380$\mu$m (see their Figure 2 for a useful visualization). The index is designed to probe the curvature of the spectrum pseudo-continuum, which varies to due absorption from water molecules. \citet{rojas-ayala12} use a custom grid of PHOENIX BT-Settl models run from the PHOENIX web simulator at. The note that the models fit the spectrum match the overall spectral shape well, although the models consistently underpredict the strength of the \ion{Na}{1} and \ion{Ca}{1} lines, even when using the highest metallicity models. \citet{rajpurohit10} note that the oscillator strengths and opacities of the \ion{Na}{1} and \ion{Ca}{1} lines may be incorrect. When restricting the spectrum to the range of the indices that they use, we find that \todo{What does our analysis yield? It should be in the table.} Because the PHOENIX grid by \citet{husser13} has more range/uniformity of parameters, we use this to fit Gl 51.

\subsection{Generic Tests}
Designed to show off what the model framework can actually do.

@@

\begin{itemize}
  \item Spectral parameters can vary by a large margin depending on which spectral library you use ($200~K$ or more).
  \item Both spectral libraries have stars that they perform better and worse on.
  \item This will be reflected in the increased level of global noise, and number of ``bad'' regions that have been instantiated.
\end{itemize}

@@

>>>>

Here we opt to take the computationally expedient approach that employs a library of model spectra, 
$\flam(\{\vt_{\ast}\}^{\rm grid})$, where $\vt_{\ast} = [T_{\rm eff}, \,\, \log{g}, \,\, Z]$ 
(in practice, the metallicity $Z$ is often parameterized by [Fe/H]).  However, it is worth noting 
that the techniques we will develop are applicable to {\it any} reasonable ``front-end" that 
generates a model spectrum.  In our adopted approach, the model spectrum for an arbitrary 
$\vt_{\ast}$ must be interpolated from among the spectral library, 
\begin{equation}
\flam(\{\vt_{\ast}\}^{\rm grid}) \leadsto \flam(\vt_{\ast}),
\label{eqn:interp}
\end{equation}
where we assign the symbol $\leadsto$ as an interpolation operator.  The multi-dimensional 
interpolation in Eq.~\ref{eqn:interp} needs to be performed many times, so computational efficiency 
is critical.  In practice, a simple tri-linear interpolation is suitably fast, but introduces an 
undesireable level of inaccuracy (particularly in the $Z$ dimension).  The interpolation quality 
can be empirically estimated by performing the operation in Eq.~\ref{eqn:interp} across a 
calculated location in $\{\vt_{\ast}\}^{\rm grid}$, and then comparing the interpolated spectrum 
with the corresponding library spectrum.  After an extensive exploration of such calculations 
\citep[see also][]{husser12}, we concluded that the best combination of speed and accuracy can be 
achieved by pre-computing a {\it refined} spectral library using a cubic spline interpolation with 
a $\{\vt_{\ast}\}^{\rm grid}$ spacing of [20\,K, 0.1\,dex, 0.1\,dex], and then performing 
tri-linear interpolation over that refined grid.  Overall, this interpolation technique is found to 
be accurate within a few percent per high resolution model pixel.  Ideally, this pre-interpolation 
could be avoided if the spectral library was computed over a refined grid (with a substantial 
up-front computational investment); but for the time being, we can empirically propagate these 
interpolation uncertainties into the likelihood calculations (as will be described in 
Section~\ref{subsec:likelihood}).  
\todo{Talk about tests to show that it is within \% accuracy}.

@@

We can further divide $\vt_{\rm obs}$ into those parameters that impact the model primarily in the 
spectral or flux dimensions; $\vt_{\rm obs} = [\vt_{\rm obs, v}, \,\, \vt_{\rm obs, f}]$.  For the 
former, we consider three kernels that contribute to the observed line-of-sight velocity 
distribution function, $\varphi_v$.  The first, $\mathcal{F}_v^{\rm inst}$,
treats the instrumental spectral broadening\footnote{For computational
efficiency, we first pre-convolve the entire raw spectral library with a slight
under-prediction of $\mathcal{F}_v^{\rm inst}$ and resample the
spectra to this lower resolution. Because the instrumental PSF acts as a
low-pass filter, this procedure can reduce the size of the library by a factor
of 10 or more while also preserving the final Fourier content of the
spectrum. Then, when a given spectrum is post-processed, the actual $\mathcal{F}_v^{\rm
inst}$ used is a correction to the original
under-prediction.}.  For illustrative purposes we assume
$\mathcal{F}_v^{\rm inst}$ is a Gaussian 
with a mean at $v = 0$ and a constant width $\sigma_v$ at all $\lambda$, although more 
sophisticated forms could be adopted.  The second, $\mathcal{F}_v^{\rm rot}$, characterizes the 
broadening induced by (projected) stellar rotation, parameterized by $v\sin{i}$ as described by 
\citet[][his Eq.~18.14]{gray08}.  And the third, $\mathcal{F}_v^{\rm dop} = \delta(v-v_z)$, 
incorporates the radial velocity through a Doppler shift.  The model spectrum is modified by the 
parameters $\vt_{{\rm obs}, v} = [\sigma_v, \,\, v\sin{i}, \,\, v_z]$ through these kernels, using 
a convolution in velocity-space,\footnote{In practice, these convolutions are performed as 
multiplications in Fourier-space to better preserve spectral information \citep[cf.,][]{tonry79}; 
the mathematical formalism is presented here for clarity.}
\begin{equation}
\flam(\vt_{\ast}, \vt_{{\rm obs}, v}) = \flam(\vt_{\ast}) \otimes \varphi_v = \flam(\vt_{\ast}) \otimes \mathcal{F}_v^{\rm inst} \otimes \mathcal{F}_v^{\rm rot} \otimes \mathcal{F}_v^{\rm dop},
\label{eqn:broadening}
\end{equation}
and then re-sampled onto the discrete wavelengths corresponding to each data pixel, 
\begin{equation}
\flam(\vt_{\ast}, \vt_{{\rm obs}, v}) \mapsto \vM(\vt_{\ast}, \vt_{{\rm obs}, v}),
\label{eqn:resampling}
\end{equation}
where the $\mapsto$ symbol denotes a re-sampling operator that maps the model spectrum onto the 
$N_{\rm pix}$-element model vector $\vM$ (and $N_{\rm pix}$ is the number of pixels in the 
spectrum).  For reference, Figure \ref{fig:broadening} shows a graphical representation of these 
post-processing steps.

@@

The inference problem as framed here has a natural hierarchical structure between the collections 
of ``interesting" parameters, $\vT = [\vt_{\ast}, \vt_{\rm obs}]$, and ``nuisance" hyperparameters, 
$\vP = [\cheb, \cov]$.  To explore the posterior probability distribution of the model conditioned 
on an observed spectrum in this case, it is convenient to use MCMC simulations with a blocked Gibbs 
sampler coupled to the Metropolis-Hastings algorithm.  In essence, this procedure works by sampling 
in a subset of parameters with Metropolis-Hastings conditioned on the current (fixed) values of the 
other parameters; after each iteration, the Gibbs sampler then updates the values of that parameter 
subset and then cycles through all the different parameter subsets in the same way (for a more 
nuanced mathematical description of the process, see Chapter 11 of \citealt{gelman13}).  The 
step-by-step prescription is as follows, where we denote the $i^{\rm th}$ iteration of the Gibbs 
sampler as a superscript index (not to be confused with the $i^{\rm th}$ pixel as in our previous 
notation):
\begin{enumerate}
\item Initialize the parameters.  One might set $\vT^0$ based on estimates in the literature or 
scaling behaviors, and make simple assumptions about $\vP^0$.  Here, we set the Chebyshev 
polynomial coefficients ($\cheb^0$) to be flat ($c_0 = 1$,  $c_{>0} = 0$ in all spectral orders) 
and assume only the trivial noise spectrum contributes to the covariance matrix (all elements of 
$\cov^0$ are zero, $\vC$ is diagonal).

@@

\item Decide whether to instantiate any additional local (non-stationary) covariance kernels.  Here 
we adopt an iterative thresholding criterion: if a local patch of the residual signal is at least 
$3\times$ the amplitude of the global covariance kernel (i.e., where $\vR \ge 3
\, a_g^i$), this will trigger the addition of a local kernel\footnote{For
practical reasons, we do not delete local kernels from the parameter list once they have been
instantiated. However, if the stellar parameters have moved to a region of
parameter space where a given local kernel is no longer required, then the
amplitude of that kernel will be driven towards $0$, the kernel would have a
negligible contribution to the covariance matrix $\vC$, and the overall behavior of the model would
be as if the kernel were deleted.}.

\item Now sample in the local covariance hyperparameters $\vp_{{\mathsf C},k}$ using one 
iteration of the Metropolis-Hastings algorithm to compute the posterior probabilities
with $\vT^i$, $\cheb^i$, and $\vp_{{\mathsf C}, g}^i$ fixed.  Update the local covariance 
hyperparameters $\vp_{{\mathsf C}, k}^{i-1} \rightarrow \vp_{{\mathsf C}, k}^i$.  
\end{enumerate}

@@

@@

\subsection{Datasets and historical analysis}
% Paragraphs as follows:
% Introduction to the two types of stars, in order to demonstrate the flexibility.
To demonstrate the flexibility of our probabilistic framework, we model spectra of two different stars acquired with different instruments. The first data set is a high resolution optical spectrum of WASP-14, an F5 star \citep{joshi09}. The second data set is a moderate resolution mid-infrared spectrum of Gl~51, an M5 star \citep{rojas-ayala12}.

% Data description of WASP-14
Our spectrum of WASP-14 was obtained with the Tillinghast Reflector Echelle Spectrograph (\emph{TRES}, \citealt{mink11}) on the Fred Laurence Whipple Observatory 1.5m telescope on June 14th, 2009. \emph{TRES} delivers a 51 echelle spectrum spanning the full optical range (3860--9100\AA) at a resolution of $R$$\sim$44,000. \citet{torres12} originally acquired the spectrum of WASP-14 to study the stellar properties of transiting exoplanet hosts. The raw spectrum was reduced using the well-tested \emph{TRES} pipeline written by L.~Buchhave and in the \ion{Mg}{0}b echelle order ($\sim$5100\AA), the spectrum has a signal to noise of 150 per resolution element.

% What other people get when they fit this star. How does our inference compare?
\citet{torres12} determined the stellar properties of WASP-14 using several different methods including {\tt MOOG}, {\tt SME}, and {\tt SPC}. While the {\tt MOOG} and {\tt SME} methods are not directly comparable to our technique because they measure individual line strengths, {\tt SPC} utilizes a moderately-sized region of spectrum and a pre-computed synthetic library that enables a direct comparison with our technique. When we make the same assumptions in \citet{torres12}, such as using the same region of spectrum (5050--5360\AA), the same spectral library, and fixing $\log g$ to the value inferred from the exoplanet transit analysis by \citet{joshi09} (amounting to a $\delta$-function prior on $\log g$$=$4.29), our inference of stellar parameters agrees to within 1$\sigma$ (see Table~\ref{table:tests}). Interestingly, when we allow $\log g$ to float, we infer a lower $T_{\rm eff}$ and $\Z$. These results demonstrate the sensitivity to assumptions about $\log g$ as well as the fact that there may exist biases in surface gravity estimates from either spectroscopy, exoplanetary transits, or both.

% Data description of Gl 51
The spectrum of Gl~51, an M5 star, was downloaded from \emph{IRTF} spectral library \citep{cushing05, rayner09}. The data was obtained with the \emph{SPEX} instrument \citep{rayner03} on the NASA \emph{IRTF}. The spectrum is a moderate resolution (R$\sim$2,000), mid-infrared spectrum from 0.7--5.5$\mu$m at signal to noise $>$400 at 2.1$\mu$m.

% What other people get when they fit this star. How does our inference compare?
To determine the stellar parameters of Gl~51, \citep{rojas-ayala12} use the equivalent widths of the 2.205$\mu$m \ion{Na}{1} doublet and 2.263$\mu$m \ion{Ca}{1} triplet along with their \HK\ index, which is defined using three passbands: 2.070--2.090$\mu$m, 2.235--2.255$\mu$m, and 2.360--2.380$\mu$m (see their Figure 2 for a useful visualization). This index is designed to probe the curvature of the spectrum pseudo-continuum, which varies to due absorption from water molecules. In order to calibrate these indices to stellar parameters, \citet{rojas-ayala12} determine these indices on a custom grid of PHOENIX BT-Settl models obtained from the PHOENIX web simulator\footnote{\url{http://phoenix.ens-lyon.fr/simulator/index.faces}}. They note that the BT-Settl models manage to fit the overall spectral shape well, but consistently under-predict the strength of the \ion{Na}{1} and \ion{Ca}{1} lines, even when using the highest metallicity models. \citet{rajpurohit10} speculate that this discrepancy may result from inaccuracies in the oscillator strengths and opacities for the \ion{Na}{1} and \ion{Ca}{1} lines. When modelling the same region of spectrum as \citet{rojas-ayala12} (2.070--2.380$\mu$m) using an updated set of PHOENIX models \citep{husser13}, we find that \todo{The results of our analysis will be in Table~\ref{table:tests}.}

\subsection{Generic Tests}

In addition to running tests under the same assumptions as historical analyses, we also allow the region kernels to play a role.

%\begin{rant about why this method rocks}
There are many advantages to placing spectroscopic inference within a Bayesian framework. Our technique uses as much as possible of the spectrum. You go to a telescope to take a spectrum, you should use the whole spectrum, not just subsets of the spectrum that are certain well chosen lines. True, these lines have been identified by the accumulated knowledge of experts to show high discriminatory power against the stellar parameters of interest, but there is a disparate gap between these indices and the core fundamental physics knowledge, since calibrations must still be derived against synthetic models. These are coming up with ``fit metrics'' that are similar to distance functions or figures of merit that would be used in approximate Bayesian computing.

These figures of merit make a lot of sense, true. But given that the raw data product from synthetic models from stars is \emph{actually a spectrum} that has been synthesized from a complicated but \emph{physical} model of a star, wouldn't it be better to relate the two through an actual \emph{likelihood function}, a physical description of the process that generates the data? In this case we use a Gaussian process to absorb the computational artifacts/inaccuracy that is a result from a \emph{imperfect} (but still extraordinarily good) stellar model. This allows us to couch the question of what parameters does the star have in a truly \emph{probabilistic} framework, meaning, we can calculate a posterior probability function of the parameters that is grounded in our physical understanding of what a star looks like. This is one of the great things about a \emph{forward-model}.

Ultimately, with the current state of physical models, generating a spectrum from a limited grid, or only being able to tweak certain parameters, the posterior that one can derive is ultimately limited by the accuracy of the spectral synthesis, including any systematic biases that may exist (and they do, given the current spread of performance of stellar libraries).

The main advantage of using a more sophisticated hierarchical likelihood function (Eqn~\ref{eqn:likelihood}) than $\chi^2$, allows the model-data mismatch to become a parameter in the model. That way there is shrinkage between estimates.

%\end{rant}

Note that using global covariance actually lets the parameters float a bit more, and provides some cushion against the tyranny of too many lines. Only so many parameters, can't infinitely tweak the model spectrum for any large region of spectrum. This parameterizes the badness of fit, better models will have a lower covariance, as evidenced by the Kurucz (-14.6) v. PHOENIX (-14.2) global cov amplitudes.

A major advantage of using the MCMC algorithm to explore the multidimensional probability space is that it provides numerical samples in each dimension. Therefore, marginalizing out a parameter (i.e., numerically integrating over a dimension in probability space) is as simple as combining all of the samples in this dimension. This enables us to present a posterior of the stellar parameters $\vT$ (Figure~\ref{fig:stellar_posterior}) which has been marginalized over all of the nuisance hyperparameters. This posterior is the final estimate of the stellar parameters which \emph{incorporates} any inherent uncertainty due to model mismatch (via the covariance hyperparameters) and flux-calibration (via the Chebyshev polynomials).

The benefit of including nuisance parameters and then marginalizing over them is that we can self-consistently model the uncertainty inherent to any spectrum while naturally capturing any degeneracy between the model parameters. If stellar parameters are estimated using a method which ignores these nuisance parameters (e.g., by-eye fitting) and the astronomer arbitrarily inflates the parameter uncertainties to reflect intuition, the degeneracy between parameters is artificially destroyed.

When iteratively fitting a spectrum, we continue to add line covariance kernels until we have covered all of the high amplitude residuals. Typically, we will add line kernels until all residuals greater than three times the amplitude of the global covariance kernel are covered.  As demonstrated by the Gibbs sampler routine, we explicitly fit for the
hyperparameters of the covariance kernels in parallel with the
stellar parameters. While including these extra parameters does increase the
dimensionality and complexity of our model, they do provide several advantages.
\emph{A priori} we do not know which regions of the spectrum are improperly
modeled. The covariance hyperparameters provide a framework to identify these
regions iteratively and in a self-consistent manner.

The traditional manner of dealing with spectral mismatch is to simply mask out the regions of the spectrum which do not agree to within a certain tolerance.  Rather than arbitrarily excluding regions of the spectrum from the fit, these regions should instead be incorporated into in the fit with the appropriate weight. A model that includes covariance is far more likely than forcing the synthetic spectrum to fit perfectly, and far more flexible than arbitrarily masking regions which do not fit. The fitting procedure allows these weights to be determined self-consistently, such that lines which are slightly wrong can still bring information to bear on the stellar parameters. In fact, for some types of stars with difficult-to-model spectra, it may be possible that the lines that are incorrect by a small amount actually provide the \emph{most information} about the stellar parameters, precisely because these lines are the most sensitive to stellar structure and consequentially are the most difficult to model correctly.

Another powerful benefit of including the covariance hyperparameters is the result of their ability to quantify and account for model-data mismatch. When a data spectrum is fit with a high-quality spectral library, the pixel residuals are less likely to be correlated, and the covariance structure will have a smaller amplitude and correlation length. This means that the inference on the stellar parameters will be more precise, in the same way that high quality data allows a more precise result. However, if the systematic mismatch between the data and model is large, then the hyperparameters will be larger and the precision of the stellar parameters will naturally inflate to respect the quality of model-data fit. The benefits of using covariance hyperparameters also extend beyond the use case of fitting a single stellar spectrum. If we fit many stars with the same set of synthetic models, we can use the structure of the covariance matrix to improve the models themselves (see \S\ref{subsec:learning}).

@@

\begin{itemize}
  \item Spectral parameters can vary by a large margin depending on which spectral library you use ($200~K$ or more). And which technique, which may be a result of choosing certain indices, certain models, or both. Therefore it is in our interest to use as wide a region of spectrum as possible.
  \item Both spectral libraries have stars that they perform better and worse on.
  \item This will be reflected in the increased level of global noise, and number of ``bad'' regions that have been instantiated.
\end{itemize}

@@

Why is it a good idea to use our likelihood function for spectroscopic inference?

By cataloguing the covariance structure of the residuals, especially those generated from strong spectral line mismatch, we collect valuable information about the quality of the synthetic spectra. After fitting many stars, the accumulated knowledge of the data-model mismatch can be used when fitting a new star. The previous structure of the covariance matrix allows us to set priors on what the covariance in certain regions of the spectrum should be, which will speed convergence for this new star. Additionally, after fitting several stars, the average value of the covariance matrix will inform us about the quality of specific spectral lines in the synthetic models.

Linking the covariance matrices of stars could be done serially, where the aggregate covariance matrix of all previous fits is used as a prior for the current iteration. Or, the covariance matrices could instead be linked hierarchically, in that the parameters describing the depth and width of a line residual for a specific star $a_1$ and $\sigma_1$, are modeled as coming from of a population of possible depths and widths for a given spectral type. Each stellar spectrum will have a slightly different realization of a spectral line, which will have some scatter about the average residual height. Linking the covariance matrix between spectra of similar stars allows us to grow more confident in our assessment that certain synthetic spectral lines are indeed outliers and should be appropriately down-weighted. In turn, as we become more certain of the weights, the stellar posteriors will become narrower and make our estimates of the stellar parameters more precise. This ability of the model to mutually inform sets of parameters is one of the major advantages of hierarchical Bayesian analysis \citep{kruschke10}.

Once determined, this average covariance matrix could be delivered to the communities that created the synthetic libraries, which would enable them to rapidly pinpoint and correct non-physical lines in the synthetic models. Alternatively, we could correct the models ourselves by using the chain of logic and mathematical post-processing that we used to created the synthetic model spectrum to reverse-engineer what the behavior of the raw synthetic spectrum \emph{should} be, at the raw $R \gtrsim$~100,000 resolution. This fundamental application of machine learning would enable us to create our own library of data driven, semi-empirical stellar models. Rather than simply assembling an empirical spectral library using only real stellar spectra, this combined approach is more powerful because the synthetic stellar atmospheres provide an actual anchor point of fundamental stellar parameters tied down by the laws of stellar physics.
